\chapter{Anarchism and Crime}
\chapterauthor{Robert Anton Wilson \& Robert Shea}

\emph{This article was first published in Green Egg, Vol. VII, No. 62, 1974.\\}

Because anarchists aim at the abolition of government, the first question they are usually asked is, ``What about murderers, thieves, rapists? The government protects us from them. Would you just let them run wild?"\\
The answer, first of all, is that government does not protect us. Its claims are a total imposture, like the fraud of a primitive shaman who claims to bring rain and warns everybody, ``If you abolish me, it will never rain again." Thus, \emph{the major crimes are all legal}; the thieves who have stolen the land and the natural resources from under our feet operate with a government franchise. These huge banks, corporations and land monopolies finance both political parties, train the corporation lawyers who become Congressmen or Presidents, and can never be successfully resisted in the courts because they own the judges, too.\\
Second, the next level of crime, the so-called Syndicate or Mafia, is also in cahoots with big government and big business, and only token arrests and light sentences are ever imposed on ``gangland leaders" --- usually rebels who have become unpopular with the higher-level mobsters. In every big city, the links are permanent and never can be under this system. The links between the national Mob and the national government are less well publicized, but books like \emph{The Politics of Opium in Southeast Asia}, the recent Harpers magazine issue on the CIA and heroin, etc., show that the heroin syndicate could not operate without high-level Federal protection.\\
Finally, the small-time free-lance criminal --- the rapist and sneak thief --- \emph{can be} arrested and prosecuted in this system; but \emph{is} he, usually? In New York, in 1972, there were 300000 burglaries but only 20000 arrests for burglary. The police are too busy protecting the high-level criminals --- as we will explain --- to have the manpower to really battle the small independents.\\
Do you deny this? Well, of course, you have been trained by the State-run schools and the mass media to deny it, do you believe your own denial? How safe do you feel in a large American city, especially after dark? Do you honestly think the government can and will protect you?

\section*{Is more law the answer?}

Many admit that they are frightened and appalled by modern American life, but they think the answer is more laws, tougher laws, an evolution toward the total Police State.\\
This is, of course, the natural direction of government. The more honest (and misguided) a politician happens to be, the more laws he will write --- to prove to himself that he is ``working" for the people. Obviously, every time the legislature meets, the honest politicians will introduce more laws, to show how hard they're working. Eventually, nothing will remain that is not covered by some law or other. Everything not compulsory will be forbidden, and everything not forbidden will be compulsory.\\
Stop and ask yourself if you really want that kind of Nazi- or Communist-style tyranny.\\
Now, even if we (or most of us) do want it --- to be protected from criminals --- and even if we escalate our progress and pass a billion new laws a year, arriving at Total Law in say five or ten years, what then? How will such a system be enforced? Kinsey estimated that to enforce our sex laws alone, 95 percent of the population would have to become either police or jail-guards --- except that they would all be in jail themselves. This is already impossible, but suppose we tried to enforce the anti-drug and anti-gambling laws, also? We would all spend our lives in Federal prisons, spending part of the day guarding others and part of the day being guarded by them.\\
This is absurd, but within the framework of government and law, how can we stop short of such a total prison-society?\\
And remember: each step in this direction --- each new law, and each new bureaucracy to enforce the new laws --- raises your tax burden. Already, you are working from January 1 to May 23 for the Federal government, to pay your IRS bill for the year. For a few months thereafter, you are working to pay nuisance taxes, state taxes, and various other concealed tawes on every item you buy, every movie you see, every drink you take. Already, it would probably be cheaper to just let yourself be robbed every week by a casual sneak-thief. Government may be more genteel than a mugger (occasionally) but it usually ends up taking more of one's money.

\section*{The function of law}

There are three kinds of laws on the books today, and to understand them is to understand the State. The first kind of law declares the State's power over you. It says: we may rob you of this much per year (taxation), we may enslave you for this period of time (the draft), we may do this and that and the other thing to you, and you cannot resist because we are your Masters. This is the earliest kind of law and was originally imposed on conquered people by conquerors. No attempt to justify it has ever been convincing to anyone bold enough to question it in the first place. It is based on mere Force; its only argument is the gun.\\
The second kind of law is coercive morality. This makes the State into an armed clergyman. It says you can enjoy yourself this way, but not that way; you can smoke this, but not that; you can drink this, but not that. Thou Shalt Not Play Parchesi On The Night of the Full Moon. Thou shalt not gamble on Sunday. Thou shalt not make love to your wife the way you and she both like, but the way the legislators like. Four million arrests a year, and an incredible expenditure of time and manpower and money, go into enforcing these laws. These are the laws that establish crimes without victims. These are the laws that everybody occasionally violates and some people violate constantly. Their only justification, as with the first type of laws, is sheer brute force. That is, without force, an man who believed in, say, the Seventh Day Adventist vegetarian diet would still obey that diet's rules; with force, the Adventists, if they get into government, can lake all of us obey it. The day is not distant wen pot-smokers will take over, and if they are vengeful, anti-booze laws will come back on the books. This stupid bullying can go on forever, each group getting its turn to impose its own prejudices on others. Anarchists say: stop it now, get off your neighbor's back, get him off your back, and let everybody enjoy his or her own lifestyle.\\
Finally, there is the third class of laws --- the class that every decent person wishes society would live by. No killing. No stealing. No rape. No fraud. Anarchists, just like you, would like to see these laws really functioning. We just don't believe that government can do that job. We think government is, always has been, and always will be preoccupied with the first two kinds of law. Read on and we will explain this.

\section*{The nature of government}

Government was instituted to guarantee that property would remain stolen. The chief function of every cop, every judge, every bureaucrat is to see that property remains stolen.\\
The first kings were conquerors. They stole the land by shot and shell, period. Then, they settled down to rob the survivors at a certain rate per year, called taxation. Next, they divided up the land among their relatives or officers in the army, who all became lords-of-the-land, landlords, and were empowered to rob the citizens at a certain other rate per year, called rent. When science and industry appeared, other satraps and sycophants of the royal families received charters to monopolize the resources and means of production, and to rob at a certain rate per year, called capital interest or profit. When banks were formed to circulate the medium of exchange (money), other charters were handed out to others in the bandit-gang, who became bank directors with a license to rob at another rate per year, called money interest or economic interest.\\
It soon became evident that those not in the gang, the majority of the population, were inclined to rob back as much as they could. The Robin Hood hero appears in all societies at this point, and most of us still admire him, although shamefacedly, since the schools and mass media tell us not to. (Still, who doesn't heroize Jesse James or John Dillinger a little?)\\
Anarchists say that the first crime was the crime of the conquerors/governors, who seized a whole land, cut it up among themselves, and proceeded to rob all of us forever by taxation, rent, corporative profit, money interest, and various sub-classes of the same basic fraud. Anarchists say that the Earth belongs to its inhabitants, not to this small ``owning" and ``governing" class of less than one percent of the population.\
Anarchists say that the way to stop crime is to stop the primordial crime, the State, and administer the land through voluntary associations (syndicates) of all the people.\\
Anarchists say that if people could work for themselves --- if they received the full product of their labor through a syndicate of fellow-workers --- almost all motivation for crime would disappear. If you didn't have to pay taxes and rent, starting tomorrow, your purchasing power would be more than doubled. If other forms of exploitation and robbery, through the financial-interest system, were also abolished, your purchasing power would more that quadruple. How much envy, how much worry about money, how much irrational fear, ulcers, nightmares, headaches and other motivations to cheat a little or steal a little would survive after this simple economic justice was achieved?

\section*{The other criminals}

``But, but --- how about the violent criminal types? How about the thrill-killers, the nuts, the psychopaths or sociopaths or sadists? How about those who simply enjoy being evil and destructive?"\\
We are not evading that question. It is absolutely necessary, however, to put it in perspective by explaining the Major Economic Crime of capitalist government (and feudal and other governments) and how other, lesser crimes mostly derive from that primordial injustice.\\
Now, after economic justice is achieved and volutary associations of all sorts (labor unions, credit unions, consumer-owned co-ops, people-owned insurance companies, rural communes, tribes, any type of free human grouping) have taken over the functions of government, \emph{some} persons, due to sickness or perversity or one damn thing or another, will still make trouble. Rape. Pilfering. Attempts to defraud. How will anarchists deal with these remaining no-goodniks?

\section*{Education and the family}

The first step in solving any social problem, like any medical problem, is prevention. Other remedies are necessary only when prevention fails.\\
Anarchists claim that the violent-nut-type of human being is produced by our current methods of child-rearing. This claim is hardly radical or extreme: every psychiatrist, every sociologist, every anthropologist, in one way or another, admits that this grave charge is true. We would not have so many rapists and other violent nuisances if our society were not, in some way, training them from birth onward to behave like that. For instance, Sweden has only a few rapes per year; the United States has one every seven minutes. One rape every seven minutes is not natural male behavior (whatever Women's Lib may say); it is a function of the sexual misery in this society.\\
Anarchists believe that the repressive, authoritarian, coercive, brutal and degrading practices currently used in the family and school are ony necessary to condition the young human to live in a government-run society. Children must be beaten or otherwise terrorized and bullied in the home and the school in order that they may ``adjust" to the terror and brutality of government as they mature. In short, a State-run society must be repressive because repression is the essence of the State.\\
Libertarian, free-form families and schools --- the open family, the Summerhill school, the free association of men, women and children without authoritarian control --- will not produce the deformed, mentally twisted, violent and ``mean" and ``crazy" types so common in our authoritarian society. So anarchists aim, first of all, to prevent violent criminals by changing the child-rearing methods that produce them.

\section*{The demoniac or monster}

There still remains the inexplicable criminal --- the guy who enjoys harming others for reasons nobody today can understand. The superstitious say he is possessed by demons; the naturalists imply that maybe he has bad genes or is a throwback to an earlier stage of evolution. Whatever the explanation, he will appear, presumably in anarchist societies, as he has appeared in all other societies, even after economic injustice and mind-warping education are abolished.\\
Human-centered societies (as distinguished from governmental or property-centered societies) have dealing with this problem for thousands of years. Tribes, clans, bands, free communes, have existed outside, before and alongside the States which get all the attention from historians. Anthropologists have investigated these free hulman groupings and have found a variety of methods of dealing with ``demoniacs," many of them as good or better that the State's traditional jails, tortures or executions.\\
Ostracism should not be underestimated. One critic of anarchism, George Orwell, actually complained that ostracism was so cruel that most people would rather fall afoul government and go to jail than be the sole ostracized person in an anarchist community.\\
Exile, widely used by governments before jail became popular, is also effective. At least, it solves the problem for the community that uses it (while, alas, passing the problem on to the unlucky community that next gets the offensive nut.)\\
The Quakers have widely practiced a form of moral forgiveness which sounds impractical to most of us, but which is murderously effective. Bertrand Russell was so impressed with this that he suggested it as a fit punishment for Stalin. Until you have seen a group of Quakers reciting somebody's sins in public, weeping over them loudly, and then forgiving and praying for the culprit, you can't imagine how much psychological impulse-to-change this generates.\\
Many anarchists believe the private defense groups are legitimate; some even are willing to allow such groups to use traditional Vigilante methods. Clarence Lee Schwartz, an American anarchist who observed this system first-hand in the old West, thought it both more humane and more effective at peace-keeping that the government law system back East. Other anarchists fear this as the possible source of a new State.\\
Most anarchists believe that criminals should not be caged under any circumstances, due to the overwhelming evidence that every prisoner comes out of a cage worse than he goes into it. Others believe, however, that punishment in a form of indemnification is compatible with libertarian ideas and should be rigorously enforced by anarchist syndicates. Under the indemnity system, every criminal must pay in cash or work or some needed good to compensate his victims (or their survivors). This certainly does the victims more good than having the criminal put in a cage and fed at community expense, to say the least of it; and is probably just as discouraging or more discouraging to every nut with even the remnant of an ability to forsee the probable results of his actions.\\
Finally, we must mention miscellaneous solutions. Just as crime in an economically just and free community will be freaky and sporadic (rather than the steady hour-after-hour terror that it is in this mad, unequal and unfree society), the remedies will also be individualized and peculiar to each situation. In some cases, undoubtedly, an anarchist community will decide the ``criminal" was right and the community was wrong; for this reason, anarchists do not believe in unalterable laws, but only in general policies.\\
The acme of anarchist theory is the principle of non-invasiveness or non-coercion --- Mind Your Own Business --- and those found to be violating this will be given, usually, some method of compensating those whose lives they have damaged. If they refuse, methods like the boycott-ostracism-exile of general cold shoulder need not always be deliberately organized against them. The good sense, the social bonds, and the sense of humor of the organic community will find some way to make them known that human tolerance, even under anarchy, is not infinite. In the Old West, men booted through town with a skunk tied around their necks, and the shoved onth the highway, often became valuable, co-operative and productive citizens in the next town, after some time to figure the likelihood of a repetition of that public amusement if they were to try similar modes of behavior again.